\propertitle

### 2019-11-17 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-17T15:08:12+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
% sample level: 279.92 nats = 1215.68 dB
% marg level: 81.525 nats =  354.06 dB
% difference in relentropies between 5 mom and 4 mom:
% sample level: 3.30882 nats = 14.37 dB
% marg level: 0.0366623 nats = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%

@@

But does the application at the full-network level lead to appreciably
different results from that at the sample level? after all we're interested
in an approximate informational sufficiency, not in an analytically exact
one. A correct answer can only be given case by case.
Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
question in the case of our first data set. The upper plot shows the
maximum-entropy distributions for a full network of $10\,000$ neurons,
constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
plot shows the maximum-entropy distributions for the sample, from two
moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
by marginalizing the full-network distribution to the sample size, from two
moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
shown. We note the following:
\begin{itemize}
\item The application to the full-network (upper plot) leads to two
  completely different distributions (even different number of modes);
  clearly the two sets of statistics are not even approximately equivalent
  for inferential purposes.
\item The application at the sample level leads to deceivingly similar
  distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
  \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
  logarithmic scale. This could lead to the erroneous conclusion that the
  two statistics are approximately equivalent.
\item The difference between marginals from the full-network distribution
  and the distributions obtained at the sample level (filled vs empty
  markers) is larger than the difference between different statistics
  (triangles vs squares).
\item The marginals of the application to the full network are closer to
  the measured frequencies than the distributions obtained at the sample
  level (although this closeness is not a valid criterion for their
  goodness).
\end{itemize}
Therefore, the maximum-entropy application at the sample level and at the
full-network level lead to different results; the latter is not only more
meaningful but also superior because it shows clearly the different
informational sufficiency of the two sets of statistics.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-17T16:12:37+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

% NUMERICAL RESULTS
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
% sample level: 279.92 nats = 1215.68 dB
% marg level: 81.525 nats =  354.06 dB
% difference in relentropies between 5 mom and 4 mom:
% sample level: 3.30882 nats = 14.37 dB
% marg level: 0.0366623 nats = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%

@@

The conclusions as regards \enquote{cooperativity} or \enquote{interaction}
that we reach by applying the maximum-entropy method to the full network,
as proposed here, are not only different from the application at the sample
level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
showing the two clearly different distributions constrained by two and four
moments, should be compared with the plot for the distributions obtained at
the sample level under the same constraints, shown in
\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
frequencies are shown as filled black triangles
(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
logarithmic scale to see the differences, which appear mainly in the tail.
\begin{figure}[!b]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

The same plot also shows the marginals obtained from applying the method to
a full network with $N=10\,000$ and the same constraints, that is, the
sample marginals of \fig~\ref{fig:plots_2_4mom}. The two-moment
distribution is represented by filled red circles
(\textcolor{myred}{\LARGE$\bm{\bullet}$}), and the four-moment one by
filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
full-network distribution fits the measured frequencies very well if we
only keep four moments.

% But does the application at the full-network level lead to appreciably
% different results from that at the sample level? after all we're interested
% in an approximate informational sufficiency, not in an analytically exact
% one. A correct answer can only be given case by case.
% Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
% question in the case of our first data set. The upper plot shows the
% maximum-entropy distributions for a full network of $10\,000$ neurons,
% constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
% from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
% plot shows the maximum-entropy distributions for the sample, from two
% moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
% (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
% by marginalizing the full-network distribution to the sample size, from two
% moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
% moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
% frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
% shown. We note the following:
% \begin{itemize}
% \item The application to the full-network (upper plot) leads to two
%   completely different distributions (even different number of modes);
%   clearly the two sets of statistics are not even approximately equivalent
%   for inferential purposes.
% \item The application at the sample level leads to deceivingly similar
%   distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
%   \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
%   logarithmic scale. This could lead to the erroneous conclusion that the
%   two statistics are approximately equivalent.
% \item The difference between marginals from the full-network distribution
%   and the distributions obtained at the sample level (filled vs empty
%   markers) is larger than the difference between different statistics
%   (triangles vs squares).
% \item The marginals of the application to the full network are closer to
%   the measured frequencies than the distributions obtained at the sample
%   level (although this closeness is not a valid criterion for their
%   goodness).
% \end{itemize}
% Therefore, the maximum-entropy application at the sample level and at the
% full-network level lead to different results; the latter is not only more
% meaningful but also superior because it shows clearly the different
% informational sufficiency of the two sets of statistics.
